### 2024-07-10 cognitive-bias.qmd
20240710 - Dunning-Krueger effect

@@

@@

@@

@@

@@

>>>>

Players' performance in fantasy football, and human behavior more generally, is complex and multiply determined.
No one can predict accurately how all players will perform.
Remain humble in your predictions; do not be more confident than you should be.

@@

As a budding empiricist, you should actively seek out information that challenges or disconfirms your beliefs and work to incorporate it into your beliefs.
Do your best to go into observation, data analysis, and data interpretation with an open mind.

@@

*Loss aversion bias* is the tendency to avoid losses rather than acquiring equivalent gains.
Loss aversion is exemplified when teams play conservatively so as "not to lose" instead of "to win."
In fantasy football, loss aversion may lead managers to start or hold onto underperforming high drafts for too long instead of a starting a more promising player out of fear of losing potential value from their initial investment.

@@

*Endowment bias* is the tendency to overvalue merely because one owns it.
For instance, a manager might overvalue a player they drafted in the first round, refusing to trade them even if they could get a better-performing player in return.

@@

The *bandwagon effect bias* is the tendency to do or believe things because other people are.
It involves social conformity.
For instance, consider if a rookie Wide Receiver has a breakout game and he is picked up in many fantasy leagues.
A given manager might pick up the player because the player is frequently being picked up in many fantasy leagues, without evaluating whether the player's success is sustainable.

@@

The *Dunning–Kruger effect bias* is the tendency for people with low ability/competency in a task to overestimate their ability.
The Dunning–Kruger effect is depicted in @fig-dunningKrueger1 and @fig-dunningKrueger2.
For instance, consider a new fantasy manager who experiences some initial wins (often called "beginner's luck").
They may attribute their successes to their skill rather than to luck.
Their [overconfidence](#sec-cognitiveBiasesOverconfidence) may lead them to believe they can win the league without much preparation.

::: {#fig-dunningKrueger1}
![](images/dunningKrueger_1.png)

Dunning–Krueger Effect. Adapted from <https://commons.wikimedia.org/wiki/File:Effet_Dunning-Kruger.svg>.
:::

::: {#fig-dunningKrueger2}
![](images/dunningKrueger_2.png)

Dunning–Krueger Effect. Adapted from <https://commons.wikimedia.org/wiki/File:Dunning-kruger_effect_-_percentile.svg>.
:::
